---
title: "large_CO-vignette"
author: "Samuel Chang"
date: "3/5/2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{large_CO-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# 1. Load the occUncertain R package and large occurrence data set for T.ornatus in Colombia
```{r}
library(occUncertain)

load("../data/T_ornatus/T_ornatus_Co.rda")

CO_observed <-
  dplyr::select(T_ornatus_Co,
    decimalLatitude,
    decimalLongitude,
    species) 
    

CO.IUCN <- IUCN.eval(CO_observed, country_map = land, exclude.area = T,  write_results = F)
large.EOO <- CO.IUCN$EOO #206994.9
large.AOO <- CO.IUCN$AOO #776

CO_observed <-
  dplyr::select(T_ornatus_Co,
    decimalLatitude,
    decimalLongitude,
    species, 
    coordinateUncertaintyInMeters)

summary(CO_observed$coordinateUncertaintyInMeters)

```

# 2. meters_to_decdeg function
```{r uncertainty from meter to radian degrees}
#NA as Zero
large_naZero <-
  meters_to_decdeg(
    T_ornatus_Co,
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    distance = "coordinateUncertaintyInMeters",
    na_action = "NA as 0"
  )

large_naZero <-
  cbind(T_ornatus_Co, large_naZero)

large_naZero <- dplyr::select(
  large_naZero,
  decimalLatitude,
  decimalLongitude,
  lon_uncertainty,
  lat_uncertainty,
  species
)

#NA as Mean
large_naMean <-
  meters_to_decdeg(
    T_ornatus_Co,
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    distance = "coordinateUncertaintyInMeters",
    na_action = "NA as mean"
  )

large_naMean <-
  cbind(T_ornatus_Co, large_naMean)

large_naMean <- dplyr::select(
  large_naMean,
  decimalLatitude,
  decimalLongitude,
  lon_uncertainty,
  lat_uncertainty,
  species
)

#NA as NA
large_naNA <-
  meters_to_decdeg(
    T_ornatus_Co,
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    distance = "coordinateUncertaintyInMeters",
    na_action = "NA as NA"
  )

large_naNA <-
  cbind(T_ornatus_Co, large_naNA)

#removes the NA cells to use function random_geo_range with uses lat/lon uncertainty
large_naNA <-
  dplyr::filter(large_naNA, !is.na(coordinateUncertaintyInMeters))

large_naNA <- dplyr::select(
  large_naNA,
  decimalLatitude,
  decimalLongitude,
  lon_uncertainty,
  lat_uncertainty,
  species
)
```

# 3. Using the random occurrences in land metrics with random_geo_range

## `NA as Mean`
```{r large_naMean_geo_range}
#1000
#----
      
large_naMean_geo_range_1000 <-
  random_geo_range(
    1000,
    large_naMean,
    taxa_col = "species",
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    lon_uncertainty = "lon_uncertainty",
    lat_uncertainty = "lat_uncertainty",
    country_map = land, 
   exclude.area = T
  )
large_naMean_geo_range_1000$NA_action <- "NA as Mean"

##----
      #100
large_naMean_geo_range_100 <-
  random_geo_range(
    100,
    large_naMean,
    taxa_col = "species",
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    lon_uncertainty = "lon_uncertainty",
    lat_uncertainty = "lat_uncertainty",
    country_map = land, 
   exclude.area = T
      )
large_naMean_geo_range_100$NA_action <- "NA as Mean"

#Z-stat
#100.EOO --> -1.950655 
(large.EOO - mean(large_naMean_geo_range_100$EOO)) / sd(large_naMean_geo_range_100$EOO)

#100.AOO --> -89.78218
(large.AOO - mean(large_naMean_geo_range_100$AOO)) / sd(large_naMean_geo_range_100$AOO)

#EOO.1K --> -1.631019
(large.EOO - mean(large_naMean_geo_range_1000$EOO)) / sd(large_naMean_geo_range_1000$EOO)

#AOO.1K --> -85.7954
(large.AOO - mean(large_naMean_geo_range_1000$AOO)) / sd(large_naMean_geo_range_1000$AOO)
```

## `Na as Zero`
```{r large_naZero_geo_range}
#1000
#----
large_naZero_geo_range_1000 <-
  random_geo_range(
    1000,
    large_naZero,
    taxa_col = "species",
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    lon_uncertainty = "lon_uncertainty",
    lat_uncertainty = "lat_uncertainty",
    country_map = land, 
   exclude.area = T
  )
large_naZero_geo_range_1000$NA_action <- "NA as Zero"

#---- 
      #100 
large_naZero_geo_range_100 <-
  random_geo_range(
    100,
    large_naZero,
    taxa_col = "species",
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    lon_uncertainty = "lon_uncertainty",
    lat_uncertainty = "lat_uncertainty",
    country_map = land, 
   exclude.area = T
  )
large_naZero_geo_range_100$NA_action <- "NA as Zero"

#Z-stat
#100.EOO -- > -0.8588903
(large.EOO - mean(large_naZero_geo_range_100$EOO)) / sd(large_naZero_geo_range_100$EOO)

#100.AOO --> -1.569794
(large.AOO - mean(large_naZero_geo_range_100$AOO)) / sd(large_naZero_geo_range_100$AOO)

#EOO.1K --> -0.9349104
(large.EOO - mean(large_naZero_geo_range_1000$EOO)) / sd(large_naZero_geo_range_1000$EOO)

#AOO.1K --> -1.564172
(large.AOO - mean(large_naZero_geo_range_1000$AOO)) / sd(large_naZero_geo_range_1000$AOO)

```

## `NA as NA`
```{r L_wiedii_low_naNA_geo_range}
#1000
#---- 
large_naNA_geo_range_1000 <-
  random_geo_range(
    1000,
    large_naNA,
    taxa_col = "species",
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    lon_uncertainty = "lon_uncertainty",
    lat_uncertainty = "lat_uncertainty",
    country_map = land, 
   exclude.area = T
  )
large_naNA_geo_range_1000$NA_action <- "NA as NA"

#---- 
      #100
large_naNA_geo_range_100 <-
  random_geo_range(
    100,
    large_naNA,
    taxa_col = "species",
    lat_col = "decimalLatitude",
    lon_col = "decimalLongitude",
    lon_uncertainty = "lon_uncertainty",
    lat_uncertainty = "lat_uncertainty",
    country_map = land, 
   exclude.area = T
  )
large_naNA_geo_range_100$NA_action <- "NA as NA"

#Z-stat
#100.EOO --> 1.727616
(large.EOO - mean(large_naNA_geo_range_100$EOO)) / sd(large_naNA_geo_range_100$EOO)

#100.AOO --> 307.84
(large.AOO - mean(large_naNA_geo_range_100$AOO)) / sd(large_naNA_geo_range_100$AOO)

#EOO.1K --> 2.805162
(large.EOO - mean(large_naNA_geo_range_1000$EOO)) / sd(large_naNA_geo_range_1000$EOO)

#AOO.1K --> 1282.771
(large.AOO - mean(large_naNA_geo_range_1000$AOO)) / sd(large_naNA_geo_range_1000$AOO)
```

##`ONly NA'
```{r}
Only <- IUCN.eval(Only_NA, country_map = land,
          exclude.area = T,
          write_results = F
        )

Only <- select()
names(large_onlyNA_geo_range_1000)[1] <- "onlyNA_EOO = 248174.8"
names(large_onlyNA_geo_range_1000)[2] <- "onlyNA_AOO = 744"
```
# Save Results as rda
```{r save iterations}
save(large_naMean_geo_range_100,
    large_naZero_geo_range_100,
    large_naNA_geo_range_100, file = "../Results/Co_results/Co_100georanges.rda")

save(naMean_geo_range_1000,
    naZero_geo_range_1000,
    naNA_geo_range_1000, file = "../Results/Co_results/Co_1000georanges.rda")

```

```{r EOO iteration plots echo=FALSE, fig.width = 10, fig.height= 15}
library(ggplot2)
library(gridExtra)
library(grid)
library(ggrepel)

theme_set(theme_gray())
theme(plot.title = element_text(hjust = 0.5))

load("../Results/Co_results/Co_100georanges.rda")

load("../Results/Co_results/Co_1000georanges.rda")

CO_1000georanges <- rbind(naMean_geo_range_1000,
                          naZero_geo_range_1000,
                          naNA_geo_range_1000)

CO_100georanges <- rbind(naMean_geo_range_100,
                         naZero_geo_range_100,
                         naNA_geo_range_100)

EOO.1000 <-
  ggplot(CO_1000georanges, aes(x = EOO, fill = as.factor(NA_action)))  +
  theme_bw() +
  scale_fill_discrete(name = "NA Parameters", guide = FALSE) +
  geom_histogram(alpha = 0.6,
                 position = 'identity') +
  geom_vline(aes(xintercept = 248174.8),
             #Only NA occ EOO
             colour = "black",
             linetype = 2) +
  geom_vline(xintercept = large.EOO, colour = "red") #observed EOO

AOO.1000 <-
  ggplot(CO_100georanges, aes(x = AOO, color = as.factor(NA_action)))  +
  theme_bw() +
  
  scale_color_discrete(name = "NA Parameters") +
  geom_histogram(alpha = 0.6,
                 position = 'identity') +
  geom_vline(aes(xintercept = 744),
             colour = "black",
             linetype = 2, ) +
  geom_vline(xintercept = large.AOO, colour = "red")

EOO.100 <-
  ggplot(CO_100georanges, aes(x = EOO, fill = as.factor(NA_action)))  +
  theme_bw() +
  scale_fill_discrete(name = "NA Parameters", guide = FALSE) +
  geom_histogram(alpha = 0.6,
                 position = 'identity') +
  geom_vline(xintercept = large.EOO, colour = "red") #observed EOO

AOO.100 <-
  ggplot(CO_100georanges, aes(x = AOO, color = as.factor(NA_action)))  +
  theme_bw() +
  
  scale_color_discrete(name = "NA Parameters") +
  geom_histogram(alpha = 0.6,
                 position = 'identity') +
  geom_vline(xintercept = large.AOO, colour = "red")

CO_plot <-
  plot_grid(
    EOO.100,
    AOO.100,
    EOO.1000,
    AOO.1000,
    ncol = 2,
    rel_widths = c(1, 1.6),
    aling = 'vh',
    labels = c("A", "B", "C", "D")
  )

legend <- get_legend(# create some space to the left of the legend
  full_plot + theme(legend.box.margin = margin(0, 0, 12)))

cowplot::ggsave(
  filename = "Results/Co_results/full_plot_grid.png",
  width = 8,
  height = 4,
  units = "in",
  dpi = 96
)