---
title: "L_wiedii_highUnc_results"
author: "Samuel Chang"
date: "6/4/2020"
output: html_document
---
```{r observed datasets with high uncertainty }
library(ConR)
load("../data/L_wiedii_highUnc.RData")

```

```{r load iteration results}
L_wiedii_high_100georanges <- read.csv("../Results/L_wiedii_high_100georanges.csv")

L_wiedii_high_1000georanges <- read.csv("../Results/L_wiedii_high_1000georanges.csv")
```

```{r #na as Mean  z-stat and hist}
#Calculate a z-statistic for the EOO of the original data set compared to the EOO values of:
(observed.EOO - mean(L_wiedii_high_100georanges$naMean_EOO)
) / sd(L_wiedii_high_100georanges$naMean_EOO)

hist(L_wiedii_high_100georanges$naZero_EOO,
  main = "L. wiedii extent of occurrence, with zero ucertainty values for NA cells")
abline(v = observed.EOO,
  col = "red")

##100 itertion is -0.5995791
(observed.EOO$EOO - mean(L_wiedii_high_1000georanges$naMean_EOO)
) / sd(L_wiedii_high_1000georanges$naZero_EOO)

hist(L_wiedii_high_1000georanges$naZero_EOO,
  main = "L. wiediee extent of occurrence, with mean ucertainty values for NA cells")
abline(v = observed.EOO$EOO,
  col = "red")

```

```{r na as Zero  z-stat and hist}
#Calculate a z-statistic for the EOO of the original data set compared to the EOO values of:

#100 iteration is -0.9961994
(observed.EOO - mean(L_wiedii_high_100georanges$naZero_EOO)
) / sd(L_wiedii_high_100georanges$naZero_EOO)

hist(L_wiedii_high_100georanges$naZero_EOO,
  main = "L. wiedii extent of occurrence, with zero ucertainty values for NA cells")
abline(v = observed.EOO,
  col = "red")

#1000 iteration is -0.8835011 --- 4551.01 sec elapsed
(observed.EOO$EOO - mean(L_wiedii_high_1000georanges$naZero_EOO)
) / sd(L_wiedii_high_1000georanges$naZero_EOO)

hist(L_wiedii_high_1000georanges$naZero_EOO,
  main = "L. wiediee extent of occurrence, with mean ucertainty values for NA cells")
abline(v = observed.EOO$EOO,
  col = "red")
```

```{r na as NA z-stat and hist}
#Calculate a z-statistic for the EOO of the original data set compared to the EOO values of:

#100 iteration <- 2161.807
(observed.EOO$EOO - mean(L_wiedii_high_100georanges$naNA_EOO)
) / sd(L_wiedii_high_100georanges$naNA_EOO)

hist(L_wiedii_high_100georanges$naNA_EOO)
abline(v = observed.EOO$EOO,
  col = "red")

#1000 iteration <- 1963.668
(observed.EOO$EOO - mean(L_wiedii_high_1000georanges$naNA_EOO)
) / sd(L_wiedii_high_1000georanges$naNA_EOO)

hist(L_wiedii_high_1000georanges$naNA_EOO)
abline(v = observed.EOO$EOO, col = "red")